Dynamic capabilities and competitive advantage

### 2019-10-14 article4_tommylars.rnw
still testint

Extant literature on DC struggles with a fundamental conception and prediction of the
theory, namely if and when DC can lead to competitive advantage. This friction has been
dubbed the "elephant in the room" (peteraf) and been tracked back to the seminal papers of
Teece and Eisenhardt, as well as their their different theoretical antecedents of
evolutionary economics (ref) and behavioral theory (ref) respectively (ref Arndt
x2). Recently, distefaneo et al propopsed a theoretical integration of these varying
conceptions of DC theory. They suggest that latent and constitutent elements of DC
represent different constructs but are integrated in that they work differently in shaping
strategic change.

DC is often critizesd for being tautological and less focus has been placed on instances
when dynamism can lead to poorer competitive outcomes. 
Structural inertia, commitment \cite{schreyogg2007}. Under what conditions are changing
capabilities counterproducttive

The relationship between dynamic capabilities and competitive advantage has been one of
great contention in the extant literature \citep{Peteraf2013,Ambrosini2009}. From this
debate a richer understanding of how different dynamic capabilities (or functional forms
of them) work differently to change different types of firm resources
\citep{Helfat2015a,Teece2007,Ambrosini2009a} with implications for competitive advantage
\citep{Peteraf2013} has emerged. Increasingly the relationship between (1) dynamic
capabilities, (2) firm resources and routines, and competitive advantage is seen as a
complex system with interactions between different types of each 1 and 2. Many conceptual
contributions have excellently discussed varieties or functional forms of either dynamic
capabilities \citep{Ambrosini2009a,Helfat2015ad,DiStefano2014} or underlying resources or
routines \citep{Winter2003,Eisenhardt2000}. However, very few, if any, papers attempt to
integrate varieties at multiple levels simultaneously. By asking if dynamic capabilities
can lead to competitive advantage and how this occurs, one is far better off asking "when"
competitive advangate is an outcome of dynamic capabiliteis and its workings.

The definition of dynamic capabilities, albeit contestst, seems to be converging on
varieties of "the ability of firms to create, modicy and extend its resource base" (ref)
\citep{Schilke2018}. Meanwhile both dynamic capabilities as the "governing entity" and
resources as "the governed entity" takes different forms in the literature (ref). Dynamic
capabilities themsleves can roughly be divided into sets of simple rules governing
undelying resources, or more complex routines developed for the same purpose
\citep{Peteraf2013}. In other words, dynamic capabilities themselves can and cannot satisfy
the VRIO condition. This contention is throughly covered in the literature (refs) and
recent contribiutions have argued for interplays between simple and complex types of
dynamic capabilities foring a more complex and initmatble dynamic system
\citep{DiStefano2014}. Moreover, granted different forms of dynamic capabilities exist
their impact on the "governed entity", the underlying firm resources, should also be
varying.

To explore this we emply two radically differnet forms of dynamic capabilities / one
geared towards transforming exitsgint resources in order to exploit them to gain
competitive advantage (refs). The other form is geraed towards seizing new opportunities in order
to explore new business opportunitieis (refs). We argue that competitive advantage can
stem from each of these, but that the mechanism varies. Dynamic capabilities geared
towards transforming for exploitation ($DC_P$) is by and of itself the result of a more
complex set of routines configured in a way that is hard to imitate because of its path
dependent and evolutionary nature. This legacy from evolutinoary economics is stressed as
an important factor for creating VRIO resources (Ref peteraf). The stream of contributions
maintaining dynamic capabilities working through the object of "changing existing
capabilities" \citep{Teece1997,Zollo2002a,Winter2003,Zahra2006,Kale2002} also maintain
that

We argue that latent and constituted dynamic capabilites behave differently and work
differnetly trhough differents modes (action or wahtever the term in DiStefano er).

{\bf LAtent = complex process routines Teece
Constitutent = simpler, ad hoc - Eisenhardt}

We explore two different modes (natures) of dynamic capabilities and how they work
differently (action) to create competitive advantage. Or is it the object of action as
either resources or opportunities?

By introducing different measures of different types of DC (latent and constiuted) as well
as two different types of action (change existing or create new) we explore the workings
of the drive train model proposed by DiStefano suggesting a fuller and more dynamic view
on how latent and consittued DC works differently.

The working of latent:
Departing from the definintion of dynamic capabilities found in Teece (ref) and subsequent
contributions in this tradition

The working of cinstuted

We identify two ditintctly different noptions of dynamic capabilities both following well
known definitions byut weith different nature (ref distefano). We show that only the
latent dc can directly and indirectly create CA,whereas the constituent element nature of
DC only works through changing lower level routines.

Different parts of the drive train works on different types of underlying routines

Looks at competing hypotheses between Eisenhardt and teexe with respect to routinization. Finds that routininzation happens at DC level less than operational level?
We argue that this depends on the purpose *action of the Dc and the composition of the
focal operational capability.

Reviewing the literature reveals that researchers have tended to identify dynamic capabilities post hoc, inferring their existence from successful organizational out- comes such as profitability and growth, as prior definitions would suggest (Table

On the other hand we have dynamic capabilites geared toward sensing and seizing new
opportunities as part of an effort of exploration. Such dynamic capabilities are described
in the literature as themselves a set of more simple routines orchestrated in a manner
such that x`

Routinization at certain level, but depends on them being outward or inward looking.

How can DC lead to competitive advangate. departing from the original definition this
happens through the modification or aquisition of firm resources. But the resopurces
themselves can and cannot be a source of competitive advantage

At a minimum, in order for something to qualify as a capability, it must work in a reliable manner’ (Helfat and Peteraf, 2003: 999). Capabilities

The bulk of contributions focus
only on heterogeneity at the dynamic capabilities level, or at

W

work dDC offers a theory of how firms achieve
competitive advantage over time. Yet the DC theory suffers from contradictory claims about
competitive advantage as its outcome. On the one hand a stream of contributions argue that
DC will lead to competitive advantage because it operates by changing operational
capabilities in face of change and hence create strategic change (ref helfat). In this
view these changes result in complex and path dependent processes that will be
idiosyncratic to the firm and hard to imitate. Consequently, they can lead to competitive
advantage.

A contrary stream of literature argues that changes of operational capabilities over time
takes the form of routines and converges towards best practice (ref EM). Such best
practices are simpler and more observable routines that are both imitable and easy to
copy. Such firm resources are then not idiosyncratic and can consequently not be the
source of competitive advantage.

The way DC work differnelty throug changing two sets of operational capabilities = complex
and path dependent processes based capabilities, and simple rules/based routines.

These contradictory positions have been thoroughly discussed in earlier work (see
\cite{Peteraf2013} for a review). On the one hand DC takes the form of simple rules
shaping underlying routines. On the other hand they take the form of complex processes.

Regarding the action and the object ofthe action, simple rules are more likely to be
involved in developing something new in response to new opportunities, while complex
rou- tines are more concerned with changing existing resources and capabilities.

In a recent "state of the art" literature review on dynamic capabilities
\cite{Schilke2018} state that much of the attention and interest surrounding the theory is
that it offer a route to competitive advantage which they argue is "the virtual Holy Grail
of strategic management" (p. 393). And indeed a considerable body of the extant literature
concerns itself with exactly this path to competitive advantage (refs from
shcilke). Still, a discourse regarding the contingencies of the dynamic
capabilities-competitive advantage (DC-CA) nexus (?) has panned out in the literature,
most notably coined "the elephant in the room" \citep{Peteraf2013}: Under which conditions
will dynamic capabilities lead to competitive advantage?

\cite{Peteraf2013} argue that the seminal papers in the dynamic capabilities theory
tradition (\cite{Teece1997} and \cite{Eisenhardt2000}) differ considerably in their view
of to what extent the DC-CA nexus exist and under what conditions it is functional. A key
tenant of this discourse is related to the level of routinization and the ability to
satisfy the VRIO condition \citep{Barney1991a} of the resource-based view
\citep{Peteraf2013}.

In this paper we aim to contribute to dissecting the elephant by asking how does the DC-CA
nexus differ between firms and what determines these differences?

We do this by exploring two paths from dynamic capabilities to competitive advantage: the
\emph{evolutionary path} and the \emph{market path}. The former is thoroughly described
and explored in the literature (refs) and stems from the core idea of the dynamic
capabilities theorizing. In this path dynamic capabilities work to change the operating
routines of the firm and hence contributing to it's competitive advantage. We argue that
this path is more suited for less dynamic environments because of its path dependent
functioning.

The \emph{market path} works not through shaping operational capabilities and routines,
but by enabling the firm to act in factor markets for new capabilities and
resources. Hence, it is less path dependent and can be more rapidly
deployed. Consequently, it is more prevalent in dynamic environments.

This paper contributes to the extant literature on the link between dynamic capabilities
and competitive advantage in three ways. First, provides an empirical test to the
drive-train model of \cite{DiStefano2014} on how dynamic capabilities can operate both
quickly and with high frequency more directly on the strategic direction of the firm by
acting in factor markets for important resources, and also more slowly through the
evolutionary path of changing operational capabilities. Second, by conceptually clarifying
the different paths to competitive advantage by linking it closely to the role of
capability evolution and the VRIO condition. Finally, we.........

CLARIFIES THE ROLE OF THE ENVIRONMENT AND MEDIATIONS HAPPENING

%% How does dynamic capabilities lead to competitive advantage?
%% Dynamic capabilities are proposed to confer a competitive advantage by adding unique value to the firm through systematic change, which enhances operational efficiency and enables increased alignment with the environment (Di Stefano et al., 2014; Peteraf et al., 2013)

%% Not all organizations possess them (Collis, 1996); their path dependencies, intangibility, complexity, and organizational specificity make them hard to imitate (Gibson & Birkinshaw, 2004; Helfat & Winter, 2011); and few other means allow organizations to purposefully change on a continuous basis (Day & Wensley, 1988; Helfat et al., 2007).

%% Consistent with the theoretical positions of Eisenhardt and Martin (2000), Helfat and Peteraf (2003), Zahra et al. (2006), and Zott (2003), among others, these researchers have argued that dynamic capabilities’ immediate purpose is to change the resource base, and that this change in the resource base, in turn, explains performance variations. According to this argument, resource changes serve as mediators through which dynamic capabilities affect performance. We

%% One of the major unresolved issues in dynamic capabilities theory is to what extent and
%% how competitive advantage is the ultimate outcome of dynamic capabilities. This very idea
%% has

%% This paper asks two questions: to what extent does a direct lnk between DC and CA exist,
%% and what determines variations in the role of mediation between them?

%% The idea that Dc impacts CA is uncontroversial (not). Little is known about how this
%% happens. The theory holds that this happens through xreating extending and modifying the
%% resource base. YEt, the change in firm capabilities can themselves lead to CA due to their
%% evlutionary nature.

%% Variations in the share of mediation stems from variation sin the environment. When envdyn
%% is high firms need to act faster and hence tends to choose strategies including resource
%% orchestration buying in factor markets cause it tkes more time to develop OR.

%% In less dynamism the role of OR increases and leads to a significant proportion of CA
%% stemming throhg the mediating effect of OR

%% There is ongoing interest in (dynamic) capabilities and how they influence firm
%% performance \citep{Fainshmidt2016a,Teece2014}. Indeed, \emph{dynamic capabilities} (DC),
%% and its front runner, the resource-based view \citep{Barney2001,Helfat2007}, is now
%% considered one of the major theories in strategic management
%% \citep{Schilke2018}. Arguably, the key reason is that this theory clearly articulates how
%% firms can alter their operating routines through dynamic capabilities. Thus, a key
%% ingredient in this theory is that DCs can alter firms operating routines, with subsequent
%% performance implications.

@@

@@

One contentious issue surrounding the DC theory is the extent to which differing views on
the nature of dynamic capabilities \citep{Peteraf2013} as well as their outcome. One
strand of contributions sees DC as routines that are simple in nature and more as best
practices to consider (EM). Thus, dynamic capabilities are more like behavioral patterns
and less idiosyncratic to the firm and will consequently not be able to satisfy the VRIO
condition for competitive advantage to emerge. Dynamic capabilities as such routines are
short lived, unstable and “substitutable” (EM: 1110) thus violating a key VRIN condition,
and they are more “more homogeneous (..) than is usually assumed” (EM: 1116). However, the
simplicity of simple rules also makes them faster do deploy and put in motion, albeit
potentially unstable. To the extent dynamic capabilities manifest themselves as simple
rules they cannot alone create competitive advantage, but could very well generate
strategic change by facilitating acquisition of resources in factor markets. One example
is the role of post-acquisition routines where routines to identify and deal with
acquisition partners are a source of generating strategic change (Eisenhardt and Sull
2001, ref). Another example is (see refs from Schilke 2018). The functioning of this
expression of dynamic capabilities are more latent and hence mainly observable ex
post. Hence we can see this as direct route to competitive advantage as this expression of
dynamic capabilities work directly in the factor market in capturing resources
needed for strategic change. We coin this direct connection as the \emph{market path}.

>>>>

%% Extant literature on dynamic capabilities is battling with a fundamental conception and
%% prediction of the theory, namely if, when and how dynamic capabilities can lead to competitive
%% advantage.

Seminal contributions to the field of dynamic capabilities have argued for both direct
\cite{Kor2005,Eriksson2014,Schilke2014a} and indirect (ref) effects of dynamic
capabilities on competitive advantage, although increasingly the bulk of the literature
focuses on the latter \citep{Schilke2018}. This indirect link is typically understood as
dynamic capabilities modifying, extending or creating firm resources (helfat) (or
lower-level operational capabilities (Winter, collis)) and thus generating competetive
advantage (Eriksson and others). Furthermore, empirical evidence keep suggesting that both direct (ref)
and indirect (ref) links exist. This warrant the question of \emph{when} direct and
indirect links between dynamic capabilities and competitive advantage is most pronounced.

Moreover, firm resources or capabiltiies themselves are far from a homogenous set and
varies among other things with respect to their uniqueness, inimatbility and value
(barney). Important contributions have thus argued for a deeper insighi into these
intermediaries and their mechanism calling for a "deeper insight into the variety of mechanisms that
underlie the performance effects of capabilities.” (Schilke, 2014a: 199). 
as well as \emph{what} different intermediaries exist.

This paper investigates the boundary conditions of the VRIO condition and how dynamic
capabilities and resources interact to create firm competitive advantage. We ask the
following research question:
Under which conditions are dynamic capabilities able to yield competietive advantage
through resource orchestration and what are the different outcomes of different types of resources.

We find thatthe dynamism surrounding hte firm represent an important boundary condition as
suggested in parts of the literature. However, DC consinutes to generate competetive
advnatage also in high-velocity environments, but doing increasingly so indirectly through
resource orchestration. Furthermore, we find that non-vrio resources are more likely to
yield competetive power in high velocity environments, and that vrio-resources tend to
maintain their CA under any condition.

@@

Under what cionditions are the direct path and indirect path most prevalant? Under which condi

The direct

@@

The link between firm resources and competitive advantage is determined by the extent to
which resources are valuable, rare, inimitable and organized (VRIO)
\citep{Barney1991a}. This principle is also used to evaluate how higher order capabilities
(e.g. dynamic capabilities \citep{Winter2003,Collis1994}. This is the core principle we
use in this article to evaluate how dynamic capabilities can and cannot create competitive
advantage for the focal firm. Furthermore, as both higher order capabilities (e.g. dynamic
capabilities) as well as lower-order resources and capabilities (operating routines and
capabilities) can and cannot satisfy the VRIO condition, we find it useful to think of a
multilevel VRIO structure in firms.

\subsection{Dynamic capabilities and VRIO}

One contentious issue surrounding the DC theory is the extent to which differing views on
the nature of dynamic capabilities \citep{Peteraf2013} as well as their outcome. One
strand of contributions sees DC as routines that are simple in nature and more as best
practices to consider (EM). Thus, dynamic capabilities are more like behavioral patterns
and less idiosyncratic to the firm and will consequently not be able to satisfy the VRIO
condition for competitive advantage to emerge. Dynamic capabilities as such routines are
short lived, unstable and “substitutable” (EM: 1110) thus violating a key VRIN condition,
and they are more “more homogeneous (..) than is usually assumed” (EM: 1116). However, the
simplicity of simple rules also makes them faster do deploy and put in motion, albeit
potentially unstable. To the extent dynamic capabilities manifest themselves as simple
rules they cannot alone create competitive advantage, but could very well generate
strategic change by facilitating the process of resource orchestration. Resource
orchestration is defined as "blablbalbla" (helfat). In other words, even if DC play out as
simple rules that are not in themselves satisfying the VRIO condition, the way they work
through creating, extending and modifying firm resources, can still yield VRIO results.

On the other hand, \cite{Teece1997} and others \cite{Helfat2007} argue that dynamic
capabiltiies will rather invariably be VRIO due to the evolutionary nature of their
development \citep{Arndt2018,Peteraf2013}. Arguing that dynamic capabilities reflect more
complex processes that co-evolve with responses to market changes over time and hence
creating idiosyncratic capabilities that are unique, valuable and hard to copy
\citep{Helfat2007,Arndt2018}. Consequently, they are likely to satisfy the VRIO condition
and thus leading to competitive advantage.

Increasingly the literature remain focused on an \emph{indirect
  path} between dynamic capabilities and competitive advantage
\citep{Barreto2010,Schilke2018}. The theory behind this path goes back to the core of
dynamic capabilities where resource orchestration happens as a response to changes in the
environment. When markets, technologies or other external factors changes, firms with high
levels of dynamic capabilities are able to seize opportunities for changes, seize these
opportunities and transform the resource base to fit the strategic change needed to stay
competitive \cite{Teece2007}. This ability to adapt is by some coined \emph{evolutionary
  fitness} defined as BLABLBALBA and stems from the ability of the firm to reconfigure
resources in the face of change. The extant literature, however, is less clear on how the
nature of the underlying resources matter in this indirect path. On the one hand we can
see

Following REF suggesting more reseaarch on dynamic capabilities should put forth
competeing hypotheses we propose these first

We introduce the notion of multilevel VRIO to integrate the interactions between different
types of capabilities at different levels. This entails that dynamic capabilities (the
high level) can consist of complex and evolutionary characteristics that are idiosyncratic
to the firm and hence VRIO by definition along the lines of how the concept is understood
by \cite{Teece1997}. On the other hand dynamic capabilities can also take the shape of
simpler routines which are short lived, unstable and substitutable thus violating a key
VRIO condition. This follows the notion of dynamic capabilities found in
h\cite{Eisenhardt2000} arguing that they are more “more homogeneous (..) than is usually
assumed” (EM: 1116).

Moreover, these dynamic capabilities are in the extant literature argued and found to work
directly \citep{Kor2005,Eriksson2014,Schilke2018} on competitive advantage which suggests
that they are indeed VRIO. On the other hand, a growing body of literature (ref) as well
a consensus on the definition of dynamic capabilities (ref), holds that they can lead to
competitive advantage by changing underlying resources and routines. In this
\emph{indirect path} the dynamic capabilities itself need not be VRIO as long as the
underlying resources themselves are.

\subsection{The role of dynamism}

The most contentious issue in the current debate on dynamic capabilities is how their
nature varies with the environment.

If htey are indeed latent

The lower level resources and routines are to be understood as conventional resources in
line with the resources-based view

WHAT IS THE ROLE OF DIRECT VS INDIRECT, HOW DOES IT RELATE TO THE ENVIRONMENT, AND WHAT
ROLE DOES THE TYPE OF UNDERLYING RESOURCE PLAY

Conventional dc literature has argued for both indirect and direct links. Indeed both
have been found, but while the indirect link is well rooted in the theory, the direct link
is to our knowledge less clear. The core idea of DC is that it works to modify underlying
resources as a response to changes in the environment and hence creating strategic change
needed to stay competitive. This resource orchestration is heavily represented in the
literature (ref) and through this process firms are able to generate CA. In such instances
the CA can stem from VRIO resources and/or VRIO DC. A VRIO DC would mean that the firm is able
to shuffle and reorganize generic resources in a way that creates a new and unique
configuration of resoures idiosyncratic and valuable to the firm. In such instances, the
outcome whould be CA. On the other hand, the DC may not by itself be VRIO but rather a set
of simple routines put in motion to change, but if the resources they reconfigure are
themselves VRIO, the outcome could just as eqsily be a competetive advantage. this
equfinaity is discussed in the literaure \cite{DiStefano2010} and makes it hard to
distingusih between the role of DC and ordinary capabilites in generating CA.

Furtemore, Considering a direct link between DC and CA, DC it self must consist of some
VRIO characteristics to garner CA. So in owther words, when is DC working directly and
indirectly_

How are dynamic capabilities working directly on competitive advantage? The microfoundations of dynamic capabilities explicated by \cite{Teece2007} is a useful
point of departure of this theoretical inquiry. He hold that for "analytical purposes,
dynamic capabilities can be disaggregated into the capacity (1) to sense and shape
opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness
through enhancing, combining, protecting, and, when necessary, reconfiguring the business
enterprise’s intangible and tangible assets". Sensing and seizing is a part of what we
would coin the "opportunity capture" of dynamic capabiliites, where as reconfiguring
concerns itself with the "resource orchestration" needed to change assets to stay
competetive.

Hence, when considering indirect and direct paths to compettetive advantage we will have
to be clear on two fronts: Are DC themselves VRIO and are the underlying resources and
routines central to the resource orchestration, VRIO.

       Path   LOW     t  HIGH     t
1 DC -> CA1 0.290 2.732 0.093 0.996
2   DC -> P 0.325 2.896 0.339 3.284
3  P -> CA1 0.064 0.462 0.171 1.913
>

The microfoundations of dynamic capabilities—the distinct skills, processes, procedures, orga- nizational structures, decision rules, and disciplines—which undergird enterprise-level sensing, seizing, and reconfiguring capacities are difficult to develop and deploy.
For analytical purposes, dynamic capabilities can be disaggregated into the capacity (1) to sense and shape opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness through enhancing, combining, protecting, and, when necessary, reconfiguring the business enter- prise’s intangible and tangible assets.

t its most fundamental level innovation research concerns itself with a simple
input-output model \citep{Adams2006} where innovation output is seen as a function of
efforts put into the innovation process. However, firms vary with respect to their ability
to utilize input and transform it into innovation output. This transformation of input
into output has been studied with different perspectives and conceptualizations such as
productivity of R\&D \citep{Gwynne2015}, innovation capability \citep{Breznik2014},
innovation factor productivity (ref) and innovation conversions rate
\citep{Duran2015}. \newline

\subsection{The idiosyncratic nature and strategic expression of IC}
The question of why firms vary in the degrees of innovation output they get from similar
levels of innovation input, leads on to a higher order capablility that coordinates and
extends the first order routines and resources as input in the innovation process. By
innovation input we mean the resources and routines put into the effort creating
innovation which composition is dealt with in a large literature (see \cite{Becheikh2006}
for a review). Innovation output means new products, processes, business models, and
organizational traits as the results of a purposeful enactment of resources (Van de Ven
1995). The mechanisms through which input is transformed to output is a reflection of firm
idiosyncratic abilities. At the meta level this concept is related to entrepreneurial
orientation \citep{Lumpkin1996}, entrepreneurial action \citep{Klein2016}, innovation
capability \citep{Breznik2014}, and absorptive capacity \citep{Cohen1990a}. We go with the
term \emph{innovation capability (IC)} defined as the "ability to continuously transform
knowledge and ideas into new products, processes and systems for the benefit of the firm
and its stakeholders"\cite[p. 384]{Lawson2001}. IC is characterized by some factors
similar for all the above mentioned concepts such as the ability to \emph{sense, seize} and
\emph{transform} opportunities into outcomes \citep{Teece2007}, and it`s idiosyncratic
nature is argued by several earlier contributions \citep{Duran2015,Klein2016}. Or as put by
\cite{Duran2015}: "While most researchers assume that a positive correlation exists
between innovation input and innovation output, the “slope” of this relationship likely
differs among organizations because of the complexity of innovation management." (ibid p
##). Innovation management (def) is the managerial levers for creating innovation
capabilities, i.e. their ability to orchestrate resources \citep{Sirmon2011} which is an
important part of what \cite{Helfat2007} calls \emph{Dynamic Managerial Capabilities
  (MDC)} defined as "the capacity of managers to purposefully create, extend, or modify
the resource base of an organization" (ibid p 24). MDC, and hence IC, is heterogeneous
between firms due to the evolutionary nature of developing such capabilities
\citep{Teece2014,Eriksson2014}. Consequently, we propose the following hypothesis:\newline

H1: IC is heterogeneous between firms due to differences in their sensing and seizing
abilities \newline

\cite{Teece2007,Teece2014} makes a sharp distinction between ordinary- and dynamic
capabilities arguing, among other dimensions, that managerial emphasis in the former
relates to cost control, whereas DC relates to entrepreneurial asset orchestration,
i.e. "doing the things right" in novel ways. A similar line of reasoning argues that
dynamic capabilities differs from ordinary routines and capabilities in that it goes
beyond best practices in operational problem solving \citep{Eisenhardt2000}. Consequently,
we would expect IC to be positively related to innovation outcomes as firms with high IC
would generate more innovation. Moreover, as managers face changing environments, the ones
with strong IC would emphasize innovation activities, whereas those with weaker IC would
have to resort to other means of adapting to change, e.g. cost efficiency. In other words,
under environmental dynamism high-IC companies will engage in sensing, seizing and
transforming opportunities, trusting the evolutionary fitness of
their dynamic capabilities, and preferring innovation as strategic choice. Meanwhile, low-IC
companies will revert to operating, administering and governing the business, trusting
their technical fitness and focus strategic choice on efficiency
\citep{Teece2014}. \newline

H2: IC is positively related to innovation activities \newline
H3: IC is negatively related to efficiency enhancing activities \newline

\subsection{The relative importance of sensing and seizing}

The concept of sensing and seizing has analogous meaning in theoretical traditions tangent
to DC. Notably entrepreneurship \citep{Hitt2011} and organizational theory
\citep{March1991}. \cite{Hitt2011} states that "entrepreneurship involves identifying and
exploiting opportunities" (ibid p 59) where identification is tied to the process of
discovery \citep{Shane2000}. Based on the definitions of opportunity discovery as DEF HERE
juxtaposed to the definition of sensing as "a scanning, creation, learning, and
interpretive activity" \citet[p. 1322]{Teece2007} we argue that they are similar in nature
and usable for exploring the importance of the latter in the innovation process. It also
bears resemblance to the concept of organizational exploration as searching, risk-taking,
discovery, experimentation, prototyping, and flexibility \citep{March1991}. Independent of
the nature of opportunities to be sensed as being either objective
\citep{Alvarez2007,Shane2000} or created \citep{Alvarez2007}, the process of sensing does
require resources and capabilities for search (ref) and a clear affinity between
opportunity and a project exist (casson and Wadeson in Mcmullen2007) suggesting s clear
inter dependency between routines and resources, and higher order capabilities of sensing
\citep{Teece2007} and making sense of the opportunity
\citep{Weick1995,March1991}. Consequently, we would expect the sensing class of dynamic
capabilities to be positively related to innovation on the one hand, and reinforcing the
impact from innovation input simultaneously.\newline

H4: Firms with strong sensing abilities will yield higher innovation output \newline

H5: Sensing as dynamic capability reinforces input in the innovation process to
generate innovation output \newline

In similar ways the concepts of seizing \citep{Teece2007} and exploitation \citep{Shane2000}
bears resemblance. However, by virtue of its nature innovation represent novel processes
within the firm so that parts of the understanding of exploitation is
redundant. Exploitation is defined as utilizing old certainties within the organization
\citep{March1991} - that is existing resources and capabilities. Innovation does by its
very nature represent a shift from existing resources and certainties (albeit often a very
incremental one, but a shift nontheless). Hence the dichotmous nature of exploration vs
exploitation does not always ring true for innovation processes. However, the discovery and
exploration of new opportunities is in any case a first step of any innovation process,
and sensing is the core class of dynamic capabiltieis in place to facilitate this
process. When the opportunity has been discovered through sensing, the firm`s ability to
capture value and generate innovation will rest on its ability to seize the opportunity -
or in other words - exploit this newly arisen opportunity. As such the exploitation
happens not on existing certainties, but of the potential future certainty arising from
the innovation opportunitiy from the sensing. Consequently, we argue that the concept of
seizing as rooted in a sense of grasping opportunities already sensed, is a more useful
construct understanding firm innovation. \newline

Similar to sensing, the process of seizing requires considerable efforts in enhancing
competencies and routines (i.e. orchestrating \citep{Hitt2011}) to grasp the opportunity
identified \citep{Shane2000}.\newline

H6: Firms with strong seizing abilities will yield higher innovation output \newline

H7: Seizing as dynamic capability reinforces input in the innovation process to
generate innovation output \newline

Despite being conceptually sequential (i.e. seizing follows sensing) their relative
importance to the innovation process is less clear from the existing literature. However,
seizing is recognizing potential from the sensed opportunities by combining novel insights
and market evolution and dynamics. Leveraging this opportunity does, however, have
implications for strategic choices and decisions to invest \citep{Teece2007,Helfat2015}
and such pursuits will likely require asset specific investments
\citep{Williamson1985}(MORE TO THE POINT REF HERE). This mechanism is fundamental as it
ties into the incentives to develop such capabilities \emph{ex ante}
\citep{Stieglitz2009}. When faced with a new venture opportunity firms will have to make
timely investment decisions with potential considerable lock in effects leading to a
Stackelberg-type investment problem \citep{Arend2009}. Comparatively, the processes of
sensing can incur with considerable search and discovery costs (REF), but "sensing does
not necessarily involve large commitments of resources, at least not relative to seizing"
\citep[p. 1343]{Teece2007}.Thus, due to its asset specific requirements, the expected
"return" from seizing should be higher than that of sensing.\newline

\noindent H8: Seizing is of relatively stronger importance in transforming innovation
input to output \newline

Finally, turning to innovation output as a first step in generating value, seizing also
plays into the process of generating business models for capturing value from innovation
\citep{Teece2007,Helfat2015}. A business model can be understood as a "pathway by which
technological innovation and knowhow combined with the utilization of tangible and
intangible assets are converted into a stream of profit"
\citep[p. 40]{Teece2017}. Consequently, seizing will play a significant role in
transforming innovation into value for the company by re-orchestrating resources and
capabilities \citep{Teece2010,Teece2017}. Hence, we will see seizing working in different
parts of the input-output process. \newline

\noindent H9: Seizing reinforces the process of generating value from innovation \newline

%\subsection{Balancing sensing and seizing - ambidexterity?}
%\begin{answer}
%  Not really sure if this is important or if it just adds confusion. Perhaps suggesting
%  writing a separate conceptual paper on the differences between exploration /exploitation
%  vs sensing/seising at a later stage
%\end{answer}

%The classical notion of the innovator`s dilemma (ref christensen) holds that
%firms face a trade-off between pursuing new avenues for business
%opportunities on the one hand (novelty), and maintaining efficient organizational
%characteristics on the other (efficiency). This trade-off is arguably leading certain
%firms to inertia in the face of environmental change due to limited abilities to pursue
%both. If this holds true we would expect a firm with considerable abilities for sensing, to be less
%able to seize, and vice versa. However, a related stream of literature argues that firms
%can do both by being \emph{ambidexterous} \citet{oreilly2004}.   \newline

\section{Studying innovation capabilities in context}
There are two important obstacles to empirically studying the working of innovation
capabilities as dynamic capabilities at the firm level. First, dynamic capabilities are
mainly considered valuable and operational in dynamic environments
\citep{Teece1997a,Helfat2007}, thus finding the proper context capturing this dynamism is
crucial. When studying the phenomena across a multitude of industries and firms with
simultaneously, the challenge arise to identify a proper context for inquiry. We
understand environmental dynamism to exhibit volatile and unpredictable characteristics
such as unstable demand conditions and shifts in industry structure, and increased
probability of supply or demand shocks \citep{Sirmon2007,Schilke2014}. In such
environments states, situations, unfamiliar to the firm occur and challenges it`s ability
to make sense of the market \citep{Weick1995}, i.e. matching the environment with a proper
internal response. Moreover, firms meeting new challenges and finding existing routines
and resource configurations to be insufficient in coping, tend to discourage truly novel
explorative solutions \citep{schreyogg2007}. This combined "matching" and "inertia" problem
will be remedied by dynamic capabilities making them a mediating factor for environmental
dynamism \citep{Schilke2014}. In addressing this challenge of context we find the 2008
global financial crisis as a useful back drop. The rapid decline in financial asset value
coupled with soaring volatility, and subsequent falling demand in the real economy,
constitute an interesting backdrop for our study \citep{Makkonen2014}.\newline

A second challenge is the measurement and operationalization of innovation capabilities as
dynamic capabilities \citep{Schilke2014}. A majority of studies are, however, relying on
proxies with single items capturing the nature of dynamic capabilities
\citep{McKelvie2009}, but several studies are increasingly building latent measures based
on observable characteristics to develop a more nuanced understanding of the phenomenon
\citep{Makkonen2014,Schilke2014}. We reconfigure the framework put forth by
\cite{Helfat2015} on how observable traits at the management level is operationalized into
sensing, seizing and transforming classes of dynamic capabilities. Following
\cite{Breznik2014} arguing for the similar nature of dynamic- and innovation capabilities,
we focus on the sensing and seizing aspect as most relevant for studying innovation in an
input-output framework. Sensing and seizing are, as argued earlier in this paper, related
to other well established constructs in the innovation and entrepreneurship literature and
hence suitable for the innovation context, most notably the idea of resource orchestration
\citep{Hitt2011} and exploration/exploitation \citep{March1991}.\newline

\cite{Helfat2015} demonstrate that certain cognitive capabilities at the management level
are antecedents of dynamic capabilities. Managerial \emph{perception} and \emph{attention}
enhance the ability for sensing. Perception relates to sensing in that management "use
cognitive frameworks they possess to “connect the dots” between changes in technology,
demographics, markets, government policies, and other factors"
\citep[p. 831]{Baron2006}. Attention, in turn, are related to perception, and hence to
sensing, as it enables managers to focus on particular information to focus "on relevant
stimuli (..)[to] facilitate environmental scanning" \citep[p. 839]{Helfat2015} necessary for
opportunity discovery \citep{Companys2007} and, consequently, sensing.\newline

Underpinning seizing are capabilities for \emph{problem solving} and
\emph{reasoning}. These are treated as one construct by \cite{Helfat2015}. In our
input-output model it is useful to distinguish between seizing innovation opportunities in
the form of deciding to invest and pursue discovered potential in the sensing process
(hence in helping generating innovation output), and the process of capturing value (i.e
generating innovation gain) through developing business models. \cite{Helfat2015} does not
explicitly make such a distinction presumably because their dependent variable is a
broader "firm performance" and because the impact of problem solving and reasoning is
working similarly in creating innovation output and innovation gain. \newline

Having found a useful context and a conceptual framework for operationalizing innovation
capabilities, we now turn to the data and methods applied in this paper.

       name Mean_est     t
3   DC -> P    0.396 6.157
4   DC -> R    0.416 8.104
5 DC -> CA1    0.239 4.018
8  P -> CA1    0.172 2.724
9  R -> CA1    0.081 1.262
> R
       Path   LOW     t  HIGH     t
1 DC -> CA1 0.417 6.585 0.144 1.513
2   DC -> R 0.388 6.078 0.534 7.525
3  R -> CA1 0.017 0.219 0.211 2.123
> P
       Path   LOW     t  HIGH     t
1 DC -> CA1 0.355 4.805 0.133 1.379
2   DC -> P 0.386 5.294 0.436 4.663
3  P -> CA1 0.155 2.193 0.232 2.315
> B
       name R_LOW_est   t.x R_HIGH_est   t.y
1 DC -> CA1     0.355 5.407      0.044 0.400
2   DC -> P     0.376 3.729      0.432 4.668
3   DC -> R     0.366 5.569      0.522 7.225
4  P -> CA1     0.158 1.989      0.220 2.357
5  R -> CA1     0.010 0.135      0.200 2.033
>

%% creating and One example
%% is the role of post-acquisition routines where routines to identify and deal with
%% acquisition partners are a source of generating strategic change (Eisenhardt and Sull
%% 2001, ref). Another example is (see refs from Schilke 2018). The functioning of this
%% expression of dynamic capabilities are more latent and hence mainly observable ex
%% post. Hence we can see this as direct route to competitive advantage as this expression of
%% dynamic capabilities work directly in the factor market in capturing resources
%% needed for strategic change. We coin this direct connection as the \emph{market path}.
